Simultaneous Elicitation of Scoring Rule and Agents Preferences for Robust Winner Determination

### 2020-01-22 draft.tex
More numbers in tables
strategies

\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \hat{\profile}, \w) - s(x; \hat{\profile}, \w) ]$ and such that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies:
\begin{align} 
\label{eq:complx}
a \pref_j x &⇔ ¬(x \pprefeq_j a)\\
\label{eq:comply}
y \pref_j a &⇔ ¬(a \pprefeq_j y) ∧ ¬((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a)).
\end{align} 
\end{claim}
\begin{proof}[Sketch of proof]
Consider our knowledge $\pprefeq_j$ about the preference of the agent $j$. 
The adversary's goal is to make the score of $y$ as high as possible and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \pprefeq_j a$); and similarly, he should put below $y$ all the alternatives he can. Two conditions must be excluded for $a$ to go below $y$. The alternatives such that $a \pprefeq_j y$ can’t be put below $y$.
Furthermore, the first objective must take priority over the second one: when an alternative should go above $x$ according to the first objective (because $¬(x \pprefeq_j a)$), and $x$ is known to be better than $y$ (thus $x \pprefeq_j y$), then $a$ should be put above $x$ (irrespective of whether $a \pprefeq_j y$), which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference of scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of scores is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
\end{proof}

@@

The \strat{Limited pessimistic} strategy uses the same criterion as the pessimistic strategy, but only considering a small set of candidate questions.
For each agent, a candidate question is chosen by considering the heuristic proposed by \citet{Lu2011}. 
For the chair, we define
\begin{align}
w^\tau & = \argmin_{\w \in W} s(\bar{y}; \bar{\profile}, \w) - s(x^{*}; \bar{\profile}, \w), \label{eq:tau} 
\end{align}
and compare component-wise $\bar{\w}$ and $\w^{\tau}$, to find the position $r$ that is most valuable to ask about (the one that maximizes $\abs{\bar{\w}_r - \w^{\tau}_r}$), and choose $\lambda$ as the middle of the interval of possible values for that rank.
%\commentOC{I am not so happy about this. If we have time, we should consider one question per rank and drop the tau business.}
We then evaluate, for each candidate question, the minimax regret for each possible answer; finally we choose the best according to $\leximax$.

Finally, the \strat{Two phases} strategy is developed in order to investigate the effect of varying the proportion of questions asked to agents and to the chair. The \strat{Two phases - vc} first asks $p$ questions to the agents, then $k-p$ questions to the chair, using in both cases Limited pessimistic to select the specific question. The \strat{Two phases - cv} strategy starts with $p$ questions to the chair, then asks to the agents. \commentOC{If our XPs exhibit no significant change between vc and cv (or if we don’t have time to run them), I will drop this distinction.}

@@

\sisetup{round-mode=places, round-precision=1, table-figures-integer=1, table-figures-decimal=1}
\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{k} & {Rnd} & {sd} & {Pes.} & {sd} & {L.\ pes.} & {sd} \\
			\midrule
			0 & 5.0 & 0 & 5.0 & 0\\
			5 & 4.958 & 0.100 & 3.741 & 0.024\\
			10 & 4.709 & 0.436 & 3.283 & 0.423\\
			15 & 4.432 & 0.467 & 2.716 &0.410 \\
			20 & 3.666 & 0.522 & 1.549 & 0.430\\
			25 & 3.068 & 0.679 & 1.424 &  0.512\\
			30 & 2.631 & 0.466 & 0.394 & 0.359\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Minimax regret in problems of size $(5, 5)$ after $k$ questions.}
	\label{fig:55}
\end{table}

\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{k} & {Rnd} & {sd} & {L.\ pes.} & {sd} \\
			\midrule
			0 & 6.0 & 0 &  &  \\
			5 &  &  &  &  \\
			10 &  &  &  &  \\
			15 &  &  &  &  \\
			20 &  &  &  &  \\
			25 &  &  &  &  \\
			30 &  &  &  &  \\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Regret in problems of size $(10, 20)$ after $k$ questions.}
	\label{fig:1020}
\end{table}

@@

The results are displayed in \cref{fig:55,fig:1020,fig:p}, where strategies are designated by Rnd for Random, 2 ph. for Two phases, Pes. for Pessimistic and L.\ pes. for Limited pessimistic.

\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{p} & {Rnd} & {sd} & {2 ph.} & {sd} \\
			\midrule
			0 &  &  &  &  \\
			20 &  &  &  &  \\
			40 &  &  &  &  \\
			60 &  &  &  &  \\
			80 &  &  &  &  \\
			100 &  &  &  &  \\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Regret in problems of size $(10, 20)$ after $100$ questions.}
	\label{fig:p}
\end{table}

>>>>

\begin{claim} \label{claim:completion}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \hat{\profile}, \w) - s(x; \hat{\profile}, \w) ]$ and such that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies:
\begin{align} 
\label{eq:complx}
a \pref_j x &⇔ ¬(x \pprefeq_j a)\\
\label{eq:comply}
y \pref_j a &⇔ ¬(a \pprefeq_j y) ∧ ¬((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a)).
\end{align} 
\end{claim}
\begin{proof}[Sketch of proof]
Consider our knowledge $\pprefeq_j$ about the preference of the agent $j$. 
The adversary's goal is to make the score of $y$ as high as possible and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \pprefeq_j a$); and similarly, he should put below $y$ all the alternatives he can. Two conditions must be excluded for $a$ to go below $y$. The alternatives such that $a \pprefeq_j y$ can’t be put below $y$.
Furthermore, the first objective must take priority over the second one: when an alternative should go above $x$ according to the first objective (because $¬(x \pprefeq_j a)$), and $x$ is known to be better than $y$ (thus $x \pprefeq_j y$), then $a$ should be put above $x$ (irrespective of whether $a \pprefeq_j y$), which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference of scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of scores is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
\end{proof}

@@

The \strat{Limited pessimistic} strategy uses the same criterion as the pessimistic strategy, but limiting it only to a small set of candidate questions. It considers the current solution of the $\MMR$, i.e. the minimax optimal alternative $x^{*}$ and the adversarial choices $\bar{y}, \bar{w}, \bar{v}$.
Then, it selects n questions (one per agent) by adopting the heuristic proposed by \citet{Lu2011}. If we consider an agent $j$, there are only three possible relations between the alternatives $x^{*}$ and $\bar{y}$ in her partial preference: either $x^{*} \ppref_j \bar{y}$, or $\bar{y} \ppref_j x^{*}$, or they are incomparable. As shown in the proof of Claim \ref{claim:completion}, in order to increase $\PMR(x^{*},\bar{y})$, an adversary should place as many alternatives as he can above $x^{*}$ in the first case, and between $\bar{y}$ and $x^{*}$ in the second. An intuitive way to reduce the pairwise max regret is, therefore, to directly ask questions involving such alternatives, hoping for an answer that would prevent the adversary to play his best game. If $x^{*}$ and $\bar{y}$ are incomparable, then we just ask the voter $j$ to compare them.
To conclude the set of candidate questions, the \strat{Limited pessimistic} strategy also selects a question to the committee. We define
\begin{align}
w^\tau & = \argmin_{\w \in W} s(\bar{y}; \bar{\profile}, \w) - s(x^{*}; \bar{\profile}, \w), \label{eq:tau} 
\end{align}
as the weight vector that minimize the $\PMR$ in the worst case.
We compare component-wise $\bar{\w}$ and $\w^{\tau}$ to find the position $r$ that is most valuable to ask about (the one that maximizes $\abs{\bar{\w}_r - \w^{\tau}_r}$), and choose $\lambda$ as the middle of the interval of possible values for that rank.
%\commentOC{I am not so happy about this. If we have time, we should consider one question per rank and drop the tau business.}
We then evaluate, for each candidate question, the minimax regret for the two possible answers; finally we choose the best according to $\leximax$.

Lastly, the \strat{Two phases} strategy is developed in order to investigate the effect of varying the proportion of questions asked to agents and to the chair. %The \strat{Two phases - vc} first asks $p$ questions to the agents, then $k-p$ questions to the chair, using in both cases Limited pessimistic to select the specific question. The \strat{Two phases - cv} strategy starts with $p$ questions to the chair, then asks to the agents. \commentOC{If our XPs exhibit no significant change between vc and cv (or if we don’t have time to run them), I will drop this distinction.}
It first asks $p$ questions to the chair, then $k-p$ questions to the agents, using in both cases \strat{Limited pessimistic} strategy to select the specific question. The aim of this strategy is to simulate the current state of the art when considering incomplete preferences. In fact, by asking enough questions to the chair we can reduce our problem into a well studied one of incompleteness of preferences knowing the voting rule. Moreover, if we invert to whom ask questions first, we fall instead, in an already studied case of uncertainty of the voting rule with a known complete profile.

@@

\sisetup{round-mode=places, round-precision=1, table-figures-integer=1, table-figures-decimal=1}
\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{k} & {Rnd} & {sd} & {Pes.} & {sd} & {L.\ pes.} & {sd} \\
			\midrule
			0 & 5.0 & 0 & 5.0 & 0 & 5.0 & 0\\
			5 & 4.958 & 0.100 & 3.741 & 0.024 & 4.350 & 0.604\\
			10 & 4.709 & 0.436 & 3.283 & 0.423 & 3.323 & 0.374\\
			15 & 4.432 & 0.467 & 2.716 &0.410 & 2.722 & 0.672\\
			20 & 3.666 & 0.522 & 1.549 & 0.430 & 2.145 & 0.716\\
			25 & 3.068 & 0.679 & 1.424 &  0.512 & 0.921 & 0.569\\
			30 & 2.631 & 0.466 & 0.394 & 0.359 & 0.470 & 0.355\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Minimax regret in problems of size $(5, 5)$ after $k$ questions.}
	\label{fig:55}
\end{table}

\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{k} & {Rnd} & {sd} & {L.\ pes.} & {sd} \\
			\midrule
			0 & 20.0 & 0 & 20.0 & 0 \\
			20 & 19.985 & 0.042 & 18.770 & 0.111 \\
			40 & 19.903 & 0.080 & 17.659 & 0.290 \\
			60 & 19.764 & 0.159 & 16.951 & 0.541 \\
			80 & 19.403 & 0.221 & 15.736 & 0.367 \\
			100 & 18.947 & 0.383 & 14.621 & 0.703 \\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Regret in problems of size $(10, 20)$ after $k$ questions.}
	\label{fig:1020}
\end{table}

@@

The results are displayed in \cref{fig:55,fig:1020,fig:p}, where strategies are designated by Rnd for Random, Pes. for Pessimistic, L.\ pes. for Limited pessimistic and 2 ph. for Two phases.

\begin{table}
	\begin{center}
		\begin{tabular}{S[table-figures-integer=2, table-number-alignment = right, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left]}
			\toprule
			{p} & {Two ph.} & {sd} \\%& {2 ph. vc} & {sd} \\
			\midrule
			0 & 100.0 & 0 \\%&  &  \\
			20 & 18.778 & 0.125 \\%&  &  \\
			40 & 17.648 & 0.473 \\%&  &  \\
			60 & 16.460 & 0.609 \\%&  &  \\
			80 & 15.470 & 0.597 \\%&  &  \\
			100 & 14.876 & 0.460 \\%&  &  \\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Regret in problems of size $(10, 20)$ after $100$ questions.}
	\label{fig:p}
\end{table}
